---
output: github_document
---
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/life cycle/#experimental)
  [![Travis build status](https://travis-ci.org/lorenzwalthert/simplificar.svg?branch=master)](https://travis-ci.org/lorenzwalthert/simplificar)
  [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/lorenzwalthert/simplificar?branch=master&svg=true)](https://ci.appveyor.com/project/lorenzwalthert/simplificar)
[![Coverage status](https://codecov.io/gh/lorenzwalthert/simplificar/branch/master/graph/badge.svg)](https://codecov.io/github/lorenzwalthert/simplificar?branch=master)
  
  <!-- README.md is generated from 
  README.Rmd. Please edit that file --> 
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%", 
  cache = FALSE
)
set.seed(1)
```


# simplificar

simplificar - for simplify in Spanish - is a high-level API for ggplot2 that 
enables fast experiments. It is envisioned as a tool for exploratory analysis 
early in the cycle of statistical model building where standardized plots 
should be available quickly, with few lines of code. The package takes 
advantage of the flexible tabular data structure implemented in the package 
tibble for handling plots after their creation.

## Installation

You can install the development version from GitHub.

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("lorenzwalthert/simplificar")
```


## Overview

The package provides two interfaces for creating plots: 

* Low-level interface: The functions `vis_[n]d_[*]()`, e.g. `vis_1d_distr()` 
  that 
  let you draw one plot at once. They have an argument `aes`, which is wrapped 
  in `ggplot2::aes_string()` and passed through to the `mapping` argument of 
  `ggplot2::ggplot()`.
  There are 
  two versions of low-level interfaces: One that outputs the plot to the 
  console (like `vis_1d_distr()` and one that writes to a file (e.g. 
  `vis_1d_distr_to_file()`) and optionally to the console too. Depending on the 
  classes of the data provided, 
  `simplificar` does a dispatch between ggplot geoms, e.g. to visualize a 
  distribution, it creates bar plots for categorical variables and density 
  plots for continuous data.
* high-level interface: The function `vis_cols()` let's you plot various plots 
  at once, specifying a `transformer`, that is, a function belonging to the 
  low-level interface introduced in the first bullet, e.g. `vis_1d_distr()`. 
  You can use tidy selectors. 

This is best understood looking at some examples.

## Examples

### High-level interface

Let's first focus on `vis_cols()`. You can create plots with the distribution 
of all variables in a data set as follows:

```{r, iris_all}
library(simplificar)
(plots <- vis_cols(iris, transformer = vis_1d_distr))
```

If visualizations should be written to files 
just use the corresponding low-level transformer (i.e. 
`vis_1d_distr_to_file`) instead. To enable console output and file output, 
set the `return_vis` to `TRUE`.
```{r, eval = FALSE}
vis_cols(iris, transformer = vis_1d_distr_to_file, return_vis = TRUE)
```


By default, all variables are selected. You can use tidy selectors (see 
`?tidyselect::vars_select_helpers()`) to only create a few plots.
```{r, iris_with}
vis_cols(iris, contains("Width"))
```

All plots are stored in the list column `gg`. We use the terminology gg table 
to refer to the tabular structure displayed above and raw gg to refer to a ggplot
in the gg list column. You can use dplyr(-like) syntax
to manipulate the gg table, e.g. you can pull out a certain raw gg. 
Let's pull the second last plot that has a numeric aesthetic.
```{r, iris_pull}
plots %>%
  dplyr::filter(class == "dbl") %>%
  pull_gg(-2)
```


You can patch different visualizations into one. All but the first argument 
passed to `merge_vis()` go into `gridExra::girid.arrange()`.
```{r, iris_merge}
plots %>%
  merge_vis(
    top = "Density plots for continous, bar charts for categorical values", 
    nrow = 2
)
```

**Transforming columns**

You can apply arbitrary transformations to one or multiple columns with 
`transform_cols()`. Your transformer need to have the vector to transform as 
first argument Further arguments to the transformer are passed at the last 
position (via `...`). Here, we can use `readr::parse_factor(x, ...)` for safe 
factor parsing.

```{r, mtcars_convert}
mtcars_converted  <- mtcars %>%
  transform_cols(c("vs", "am"), transformer = "readr::parse_factor", levels = 1:0) %>%
  transform_cols(c("cyl"), transformer = "readr::parse_factor", levels = c(4, 6, 8))

```

**Internal dispatch for different classes of data**

For ggplot2, it is essential to use the correct class for each variable, 
otherwise, the plot may not look as expected. That is why `simplificar`
offers an automatic dispatch layer. Assuming you want to create scatter plots. 
You can select the corresponding transformer `vis_2d_point`. 
`simplificar` will check if any plot you draw has categorical variables only. 
If so, you probably want to use `ggplot2::geom_jitter()` instead of 
`ggplot::geom_point()`. `simplificar` will take care of that and select the 
geom according to the variable class.

```{r, mtcars_converted_vis}
multiple_vis <- mtcars_converted %>%
  vis_cols(vs, "cyl", transformer = vis_2d_point) 

multiple_vis %>%
  merge_vis()
```

If a visualization has multiple aesthetics, each of them is stored in a separate
element in the list column `aes` of the gg table (see below). The same is true 
for the class attribute. The columns `aes_string` and `class_string` contain all
classes and aesthetics pasted together.

```{r}
multiple_vis
```

**Generating all pair-wise point visualizations**

If you supply more variables to `vis_cols()` than the indicated 
transformer has dimensions, it simply creates all combinations. This is really useful if 
you want to create many plots.

```{r, mtcars_converted_merged_jitter}
mtcars_converted %>%
  vis_cols(vs, contains("hp"), "cyl", transformer = vis_2d_point) %>%
  merge_vis(ncol = 3)
```


**Generating all pair-wise distribution visualizations**

We can visualize the distribution in a similar fashion. Note that we can also 
use the transformer `vis_distr` and specify the number of dimensions in the plot
manually via `k_dimensional` instead of using `vis_2d_distr`.

```{r, mtcars_converted_distr}
mtcars_converted %>%
  vis_cols(vs, contains("hp"), "mpg", transformer = vis_distr, k_dimensional = 2) %>%
  dplyr::slice(-1) %>%
  mutate_gg(ggplot2::geom_point()) %>%
  merge_vis()
```

We use `mutate_gg()` (see below) to add the raw data points to the plots.

### Low-level interface

If you need more control over the visualizations you create, you can use the 
low-level interface.

**Manipulating the geom**

For example, in the above plot in the middle, you may don't want the jitter
effect to be so strong. Therefore, use the transformer directly and pass 
additional arguments that should go into the ggplot geom (in our
case `ggplot2::gemo_jitter()`) via `...`.

```{r, low_level_intro_jitter}
vis_2d_point(mtcars_converted, c("vs", "cyl"), width = 0.1, height = 0.1) %>%
  pull_gg()
```

We can also override the geom determined by the internal dispatch of `simplicar`
by specifying the geom argument ourself. Hence, we can use the initial `mtcars` 
data set again and we don't nee to rely on variable class conversion to jitter 
the points. 
```{r disabled_geom_dispatch_jitter}
# let's override the geom dispatch
disabled_geom_dispatch <- vis_2d_point(mtcars, c("vs", "cyl"), 
  geom = ggplot2::geom_jitter, width = 0.1, height = 0.1) %>%
  pull_gg()
disabled_geom_dispatch
```

**using ggplot2 aritmetric additions**

The way the axis are labeled in the above plot is a bit unfortunate. Recall 
that `pull_gg()` returns a normal ggplot, so you can use the `+` operator to 
customize it further.

```{r, low_level_manipulate}
disabled_geom_dispatch + 
  ggplot2::scale_x_continuous(breaks = c(0, 1))

```

You can also use `mutate_gg()` to manipulate certain raw gg objects.
As stated above, you can use standard `ggplot2` to modify raw ggplots.
`mutate_gg()` takes a gg table, an addition you want to make plus the
row numbers in the gg table you want to add the addition operation.
Below, we add a mean line to plot 1 and 2.

```{r, mutate_gg}
mtcars %>%
  vis_cols(vs, contains("hp"), "cyl", transformer = vis_2d_point) %>%
  mutate_gg(ggplot2::stat_summary(fun.y = mean, geom = "line"), 1, 2) %>%
  merge_vis(ncol = 3)
```

If you don't pass any value to `...`, all columns are modified.

Note that you can also use `purrr::partial(..., .first = FALSE)` and per-fill 
some arguments of a low-level interface function and then feed the new function into 
the high-level interface as your *adjusted* transformer. Make sure you set 
`.first = FALSE`.
```{r, purrr_partial}
vis_2d_point_with_weak_jitter <- purrr::partial(vis_2d_point, 
  width = 0.1, height = 0.1, .first = FALSE
)
vis_cols(mtcars_converted, vs, cyl, hp,
  transformer = vis_2d_point_with_weak_jitter
)
```

Note that you the you can only pre-fill arguments that are not determined by 
`vis_cols()`, i.e. you cant' set `aes` and `names`.
